---
title: "Team_Essay_4"
author: "Sahil, Ariel, Ilya, Lekha, Anthony, Trang"
date: "4/4/2021"
output: pdf_document
---

## Introduction

In this essay, our team will predict the species of a flower using the K Nearest Neighbors (KNN) algorithm. Given the sepal length, sepal width, petal length, and petal width in centimeters, we will classify the flower as either an Iris-setosa (ISE), Iris-versicolor (IVE), or Iris-virginica (IVI). Our dataset, flowers, contains 50 flowers from each of 3 species of iris. 

## Formula and Basics

Anthony

## Loading required R packages

```{r}
library("readxl")
library("class")
library("tidyverse")
library("gmodels")
library("flexplot")
```

Anthony and Ilya.

## Data description

Anthony

## Examples of data and problem

Anthony

## Visualization

```{r}
flowers <- read_excel("IRIS.xlsx")

table(flowers$species) # It gives the number of indiviudal species

flowers$species <- factor(flowers$species, levels = c("Iris-setosa", "Iris-versicolor", "Iris-virginica"), labels = c("ISE", "IVE", "IVI"))

my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")  
# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}
# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 19, col = my_cols[flowers$species])
}
# Create the plots
pairs(flowers[,1:4], 
      lower.panel = panel.cor,
      upper.panel = upper.panel)
```
The scatter plot matrix above graphs all the pairs of predictor dimensions with the blue points representing the ISE species, the yellow points representing the IVE species, and the red points representing the IVI species. We observe that petal width and petal length are highly correlated with a value of R = 0.96 as well as sepal length and petal length with a value of R = 0.87.

## Analysis

## Computation 

```{r}
round(prop.table(table(flowers$species)) * 100, digits = 1)  # It gives the result in the percentage form rounded of to 1 decimal place( and so it’s digits = 1)

# Normalize the data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

flowers_n <- as.data.frame(lapply(flowers[1:4], normalize))
flowers_2 <- flowers_n[-5]
flowers_2
```

## Interpretation

```{r}
summary(flowers_n$sepal_length)
summary(flowers_n$sepal_width)
summary(flowers_n$petal_length)
summary(flowers_n$petal_width)
```


## Model Evaluation

```{r}
# Divide the data-set into training and testing sets.
flowers_train <- flowers_n[1:90, ]
flowers_test <- flowers_n[91:150, ]

flowers_train_labels <- flowers_2[1:90, ]
flowers_test_labels <- flowers_2[91:150, ] # This code takes the species factor in column 1 of the flowers data frame and on turn creates flowers_train_labels and flowers_test_labels data frame.
```

## Regression Line

Ariel.

## Model Assessment

```{r}
length(flowers_train)
length(flowers_train_labels)
# Error to figure out on Tuesday.
flowers_test_pred <- knn(train = flowers_train, test = flowers_test, cl = flowers_train_labels, k=12)

CrossTable(x = flowers_test_labels, y = flowers_test_pred, prop.chisq = FALSE)
```

## Model Summary

Sahil and Ariel.

## Prediction and Model Accuracy

Sahil and Ariel.

## Conclusion

Lekha and Trang.

## Reference

Lekha and Trang.