---
title: "Team_Essay_4"
author: "Sahil, Ariel, Ilya, Lekha, Anthony, Trang"
date: "4/4/2021"
output: pdf_document
---

## Introduction

Anthony and Ilya.

## Formula and Basics

Anthony and Ilya.

## Loading required R packages

```{r}
library("readxl")
library("class")
library("tidyverse")
library("gmodels")
```

Anthony and Ilya.

## Data description

Anthony and Ilya.

## Examples of data and problem

Anthony and Ilya.

## Visualization

Anthony and Ilya.

## Analysis

## Computation 

```{r}
flowers <- read_excel("IRIS.xlsx")

table(flowers$species) # It gives the number of indiviudal species

flowers$species <- factor(flowers$species, levels = c("Iris-setosa", "Iris-versicolor", "Iris-virginica"), labels = c("ISE", "IVE", "IVI"))

round(prop.table(table(flowers$species)) * 100, digits = 1)  # It gives the result in the percentage form rounded of to 1 decimal place( and so it’s digits = 1)

# Normalize the data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

flowers_n <- as.data.frame(lapply(flowers[1:4], normalize))
flowers_2 <- flowers[-5]
flowers_2
```

## Interpretation

```{r}
summary(flowers_n$sepal_length)
summary(flowers_n$sepal_width)
summary(flowers_n$petal_length)
summary(flowers_n$petal_width)
```


## Model Evaluation

```{r}
# Divide the data-set into training and testing sets.
flowers_train <- flowers_n[1:90, ]
flowers_test <- flowers_n[91:150, ]

flowers_train_labels <- flowers_2[1:90, ]
flowers_test_labels <- flowers_2[91:150, ] # This code takes the species factor in column 1 of the flowers data frame and on turn creates flowers_train_labels and flowers_test_labels data frame.
```

## Regression Line

Ariel.

## Model Assessment

```{r}
length(flowers_train)
length(flowers_train_labels)
# Error to figure out on Tuesday.
flowers_test_pred <- knn(train = flowers_train, test = flowers_test, cl = flowers_train_labels, k=12)

CrossTable(x = flowers_test_labels, y = flowers_test_pred, prop.chisq = FALSE)
```

## Model Summary

Sahil and Ariel.

## Prediction and Model Accuracy

Sahil and Ariel.

## Conclusion

Lekha and Trang.

## Reference

Lekha and Trang.